---
title: "survival analysis"
output:
  workflowr::wflow_html:
    code_folding: hide
    toc: true
date: "2025-05-06"
---


```{r, echo=F, message=F, warning=F, results=F}
library(rprojroot)
root <- rprojroot::find_rstudio_root_file()
source(file.path(root, "../../AllCollaboration/Collaborations/analysis/Rfunctions.R"))
root <- rprojroot::find_rstudio_root_file()
```







```{r, message=F, warning=F, eval=F}

path="C:\\Shengtong\\Research\\AllCollaboration\\2024\\202405\\Pradeep\\DDWI\\DDWI\\"
all_data=readRDS(paste(path, "all_data.rds", sep=""))


# Loop over each element of all_data and process demographic and claims
processed_list <- lapply(all_data, function(data_item) {
  # Standardize column names
  colnames(data_item$`Demographic (1)`) <- colnames(data_item$`Demographic `)
  colnames(data_item$`Claims and Provider (1)`) <- colnames(data_item$`Claims and Provider `)
  
  # Combine
  demographic_combined <- rbind(data_item$`Demographic `, data_item$`Demographic (1)`)
  claims_combined      <- rbind(data_item$`Claims and Provider `, data_item$`Claims and Provider (1)`)
  
  # Remove unnecessary columns
  demographic_combined <- demographic_combined %>% 
    select(-`Subscriber Address Line 1`, -`Subscriber Address Line 2`, -`Subscriber City`, -`Subscriber Zip Code`)
  
  claims_combined <- claims_combined %>% 
    select(-`Allowed Amount`, -`Paid Amount`, -`Provider Name`, 
           -`Clinic Address Line 1`, -`Clinic Address Line 2`, 
           -`Clinic City`, -`Clinic Zip Code`)
  
  # Return the result for this element
  list(
    demographic = demographic_combined,
    claims = claims_combined
  )
})

# Now combine all demographic and claims across all 10 elements
demographic_combined_all <- do.call(rbind, lapply(processed_list, `[[`, "demographic"))
claims_combined_all      <- do.call(rbind, lapply(processed_list, `[[`, "claims"))
claims_combined_all$member_tooth_number=paste(claims_combined_all$`Unique Member ID`, claims_combined_all$`Tooth Number`, sep=":")

library(qs)

qsave(demographic_combined_all, paste(path, "demographic_combined_all.qs", sep=""))
qsave(claims_combined_all, paste(path, "claims_combined_all.qs", sep=""))
```



```{r, message=F, warning=F}
library(qs)
path="C:\\Shengtong\\Research\\AllCollaboration\\2024\\202405\\Pradeep\\DDWI\\DDWI\\"
demo <- qread(paste(path, "demographic_combined_all.qs", sep=""))
claims <- qread(paste(path, "claims_combined_all.qs", sep=""))
```


```{r, message=F, warning=F, eval=T}

compute_survival <- function(tooth_id, claims, demo, index = NA, start_code) {
  message("Processing [", index, "/", length(teeth_sample), "]: ", tooth_id)

  # Same logic as before...
  single_tooth_claim_data <- claims %>%
    filter(member_tooth_number == tooth_id)

  if (nrow(single_tooth_claim_data) == 0) {
    return(data.frame(
      Tooth_ID = tooth_id,
      Start_Date = NA,
      End_Date = NA,
      Survival_Time_Days = NA,
      Event_Observed = NA,
      Index = index,
      provider_specialty=NA,
      surface_area_code=NA
    ))
  }

  member_id <- single_tooth_claim_data$`Unique Member ID`[1]

  single_tooth_demo_data <- demo %>%
    filter(`Unique Member ID` == member_id)

  single_tooth_claim_data$`Service Date` <- as.Date(single_tooth_claim_data$`Service Date`)
  termination_date <- as.Date(single_tooth_demo_data$`Member Termination Date`[1])

  start_row <- single_tooth_claim_data %>%
    filter(`Submitted Procedure Code` == start_code) %>%
    slice(1)

  if (nrow(start_row) == 0) {
    return(data.frame(
      Tooth_ID = tooth_id,
      Start_Date = NA,
      End_Date = NA,
      Survival_Time_Days = NA,
      Event_Observed = NA,
      Index = index, 
      provider_specialty=NA,
      surface_area_code=NA
    ))
  }

  start_date <- start_row$`Service Date`[1]

  has_7140 <- single_tooth_claim_data %>%
    filter(`Submitted Procedure Code` == 7140)

  if (nrow(has_7140) > 0) {
    end_date <- min(has_7140$`Service Date`, na.rm = TRUE)
    event_observed <- 1
  } else {
    end_date <- termination_date
    event_observed <- 0
  }

  survival_time <- as.numeric(difftime(end_date, start_date, units = "days"))
  
  provider_specialty=claims %>%
    filter(member_tooth_number == tooth_id) %>%
    select(`Provider Specialty Name`) %>%
    pull()
  surface_area_code=claims %>%
    filter(member_tooth_number == tooth_id) %>%
    select(`All Surfaces Area Code`) %>%
    pull()
  
  

  return(data.frame(
    Tooth_ID = tooth_id,
    Start_Date = start_date,
    End_Date = end_date,
    Survival_Time_Days = survival_time,
    Event_Observed = event_observed,
    Index = index, 
    provider_specialty=provider_specialty[1], 
    surface_area_code=surface_area_code[1]
    
  ))
}

```


* start with code 3120/3220, for each tooth, find out the follow up codes in the enrollment period, it's treated as a failure if code 7140 appears before the enrollment expires, and censored otherwise. 






## D3220 Therapeutic Pulpotomy













```{r, message=F, warning=F, eval=F}
teeth_sample=claims %>% filter(`Submitted Procedure Code`==3220) %>% select(member_tooth_number) %>% pull()


survival_results <- map2_dfr(
  .x = teeth_sample,
  .y = seq_along(teeth_sample),
  .f = ~compute_survival(tooth_id = .x, claims = claims, demo = demo, index = .y, start_code=3220)
)
#qsave(survival_results, paste(path, "survival_results_for_code3220.qs", sep=""))
```

### survival analysis 



```{r, message=F, warning=F}
survival_results=qread(paste(path, "survival_results_for_code3220.qs", sep=""))
# Assuming your data frame is named survival_results
survival_results_clean <- survival_results[!grepl("^9999", survival_results$End_Date), ]


table(survival_results_clean$provider_specialty)
```



```{r, message=F, warning=F}

library(survival)
library(survminer)

# Create the survival object
surv_obj <- Surv(time = survival_results_clean$Survival_Time_Days,
                 event = survival_results_clean$Event_Observed)

# Fit the Kaplan-Meier model
km_fit <- survfit(surv_obj ~ 1, data = survival_results_clean)

# Plot the KM curve: red curve, gray confidence interval
ggsurvplot(km_fit,
           data = survival_results_clean,
           conf.int = TRUE,
           #conf.int.fill = "gray",       # Set confidence band to gray
           censor.shape = NA,      # remove vertical censoring marks
           xlab = "Time (days)",
           ylab = "Survival probability",
           title = "Kaplan-Meier Curve for Tooth Survival",
           palette = "red")              # Set curve color to red
```







```{r, message=F, warning=F}

fit <- survfit(Surv(Survival_Time_Days, Event_Observed) ~ provider_specialty, data = survival_results_clean)

# Extract labels in the correct order
labels <- sub(".*=", "", names(fit$strata))

ggsurvplot(fit,
           data = survival_results_clean,
           pval = TRUE,
           risk.table = TRUE,
           xlab = "Time (days)",
           ylab = "Survival Probability",
           censor.shape = NA,      # remove vertical censoring marks
           legend.title = "",
           legend.labs = labels,
           palette = "Dark2",
           risk.table.fontsize = 2,
           tables.theme = theme_bw(base_size = 8))
```


### permanent (1-32) vs  primary tooth (A-T)


```{r, message=F, warning=F}

# Extract the tooth code (part after :)
survival_combined_clean <- survival_results_clean %>%
  mutate(
    Tooth_Code = sub(".*:", "", Tooth_ID),
    Tooth_Type = ifelse(Tooth_Code %in% as.character(1:32), "Permanent",
                        ifelse(Tooth_Code %in% LETTERS[1:20], "Primary", NA))
  )

# Check distribution
table(survival_combined_clean$Tooth_Type, useNA = "ifany")

# Build survival object
surv_obj <- Surv(
  time = survival_combined_clean$Survival_Time_Days,
  event = survival_combined_clean$Event_Observed
)

# Fit Kaplan-Meier by tooth type
fit <- survfit(surv_obj ~ Tooth_Type, data = survival_combined_clean)

# Plot survival curves
ggsurvplot(
  fit,
  data = survival_combined_clean,
  risk.table = TRUE,
  pval = TRUE,
  conf.int = TRUE,
  censor.shape = NA,      # remove vertical censoring marks
  legend.title = "Tooth Type",
  legend.labs = c("Permanent", "Primary"),
  xlab = "Days",
  ylab = "Survival Probability"
)

```



## D3120 Pulp cap indirect



```{r, message=F, warning=F, eval=F}
start_code=3120
teeth_sample=claims %>% filter(`Submitted Procedure Code`==start_code) %>% select(member_tooth_number) %>% pull()
#teeth_sample=teeth_sample[1:3]


survival_results <- map2_dfr(
  .x = teeth_sample,
  .y = seq_along(teeth_sample),
  .f = ~compute_survival(tooth_id = .x, claims = claims, demo = demo, index = .y, start_code=3120)
)
#qsave(survival_results, paste(path, "survival_results_for_code", start_code, ".qs", sep=""))
```






```{r, message=F, warning=F}
start_code=3120
survival_results=qread(paste(path, "survival_results_for_code", start_code, ".qs", sep=""))
# Assuming your data frame is named survival_results
survival_results_clean <- survival_results[!grepl("^9999", survival_results$End_Date), ]
table(survival_results_clean$provider_specialty)
```




```{r, message=F, warning=F}

library(survival)
library(survminer)

# Create the survival object
surv_obj <- Surv(time = survival_results_clean$Survival_Time_Days,
                 event = survival_results_clean$Event_Observed)

# Fit the Kaplan-Meier model
km_fit <- survfit(surv_obj ~ 1, data = survival_results_clean)

# Plot the KM curve: red curve, gray confidence interval
ggsurvplot(km_fit,
           data = survival_results_clean,
           conf.int = TRUE,
          # conf.int.fill = "gray",       # Set confidence band to gray
           censor.shape = NA,      # remove vertical censoring marks
           xlab = "Time (days)",
           ylab = "Survival probability",
           title = "Kaplan-Meier Curve for Tooth Survival",
           palette = "red")              # Set curve color to red
```



```{r, message=F, warning=F}

fit <- survfit(Surv(Survival_Time_Days, Event_Observed) ~ provider_specialty, data = survival_results_clean)

# Extract labels in the correct order
labels <- sub(".*=", "", names(fit$strata))

ggsurvplot(fit,
           data = survival_results_clean,
           pval = TRUE,
           risk.table = TRUE,
            censor.shape = NA,      # remove vertical censoring marks
           conf.int = TRUE,
           xlab = "Time (days)",
           ylab = "Survival Probability",
           legend.title = "",
           legend.labs = labels,
           palette = "Dark2",
           risk.table.fontsize = 2,
           tables.theme = theme_bw(base_size = 8))
```



### permanent (1-32) vs  primary tooth (A-T)


```{r, message=F, warning=F}

# Extract the tooth code (part after :)
survival_combined_clean <- survival_results_clean  %>%
  mutate(
    Tooth_Code = sub(".*:", "", Tooth_ID),
    Tooth_Type = ifelse(Tooth_Code %in% as.character(1:32), "Permanent",
                        ifelse(Tooth_Code %in% LETTERS[1:20], "Primary", NA))
  )

# Check distribution
table(survival_combined_clean$Tooth_Type, useNA = "ifany")

# Build survival object
surv_obj <- Surv(
  time = survival_combined_clean$Survival_Time_Days,
  event = survival_combined_clean$Event_Observed
)

# Fit Kaplan-Meier by tooth type
fit <- survfit(surv_obj ~ Tooth_Type, data = survival_combined_clean)

# Plot survival curves
ggsurvplot(
  fit,
  data = survival_combined_clean,
  risk.table = TRUE,
  pval = TRUE,
  conf.int = TRUE,
  censor.shape = NA,      # remove vertical censoring marks
  legend.title = "Tooth Type",
  legend.labs = c("Permanent", "Primary"),
  xlab = "Days",
  ylab = "Survival Probability"
)

```



## 3220 vs 3120 


```{r, message=F, warning=F}
survival_results1=qread(paste(path, "survival_results_for_code", 3220, ".qs", sep=""))
survival_results2=qread(paste(path, "survival_results_for_code", 3120, ".qs", sep=""))
survival_combined=rbind(survival_results1, survival_results2) 
survival_combined$code=c(rep("3220", nrow(survival_results1)), rep("3120", nrow(survival_results2)))
survival_combined_clean <- survival_combined[!grepl("^9999", survival_combined$End_Date), ]
```





```{r, message=F, warning=F}
# Create a Surv object
surv_object <- Surv(time = survival_combined_clean$Survival_Time_Days, 
                    event = survival_combined_clean$Event_Observed)

# Fit a survival model by 'code'
fit <- survfit(surv_object ~ code, data = survival_combined_clean)

# Plot the survival curves
ggsurvplot(
  fit, 
  data = survival_combined_clean,
  pval = TRUE,                # Add p-value for log-rank test
  conf.int = TRUE,            # Add confidence intervals
  censor.shape = NA,      # remove vertical censoring marks
  risk.table = TRUE,          # Show number at risk
  legend.title = "Code",      
  xlab = "Time in Days",      
  ylab = "Survival Probability",
  palette = "Dark2",           # Use a nice color palette
  risk.table.fontsize = 3    # smaller font size (try 2 to 4 for smaller text)
)
```


* small p value suggests the survival curves differ significantly between code 3120 and 3220. 