---
title: "analysis for year 2019"
output:
  workflowr::wflow_html:
    code_folding: hide
    toc: true
date: "2024-10-09"
---

```{r, echo=F, message=F, warning=F}
source("C:/Shengtong/Research/AllCollaboration/Collaborations/analysis/Rfunctions.R")
```



## 1. among people enrolled in HMO/FFS, how many went to dentist at least once


```{r, message=F, warning=F}

#data_2019$demographics_2019 %>% filter(demographic_2019_check=="Yes") %>% nrow()  # 1612620
#length(unique(data_year2019$hmo_claims_2019$RECIPIENT_ID)) # 56763
#length(unique(data_year2019$ffs_claims_2019$RECIPIENT_ID)) # 59361

ggplot(data.frame(source=c( "Demographs", "HMO claims", "FFS claims"), num=c(1612620, 56763, 59361)), aes(x=source, y=num, fill=source))+ 
geom_bar(position = "dodge",stat="identity")+
   xlab("")+ylab("No. unique ids")+
 #  ylim(c(0,1))+
   theme(axis.text.x = element_text(angle =0, vjust = 0.5, size=10))+
   ggtitle("year 2019")+
   geom_text(aes(label=num), position=position_dodge(width=0.9), vjust=-0.25)+  # add numbers over bars 
  #geom_signif(annotations ="p=0.074", y_position = 0.52  ,xmin=1, xmax=2, size=1, vjust=-0.5, textsize=3.5, color="black")+ # textsize is for text over the bar; size is for width of lines 
  # geom_signif(annotations ="p<0.001", y_position = 0.59  ,xmin=1, xmax=3, size=1, vjust=-0.5, textsize=3.5, color="red")+
  #geom_signif(annotations ="p<0.001", y_position = 0.86  ,xmin=1, xmax=4, size=1, vjust=-0.5, textsize=3.5, color="red")+
   theme(plot.title = element_text(hjust = 0.5, size=15))+  #center the title
  theme(legend.title=element_blank())+
  theme(legend.position = "none")
```

* details about how these numbers are calculated  are [here](https://han16.github.io/Oral-public-health/part2.html)
* FFS: `59361/1612620
[1] 0.03681028`; HMO: `56763/1612620
[1] 0.03519924`

* total number from `Demographs`. 



```{r, message=F, warning=F}
FFS_count=c(393420, 385650, 388230, 389031, 384248, 384563, 387879, 387891, 387930, 390666, 387225, 389706)
HMO_count=c(803755, 813129, 811751, 812222, 818355, 818552, 815305,	817917,	818435,	814588,	816982,	814091)


ggplot(data.frame(source=c( "FFS_enrolled", "HMO_enrolled", "HMO claims", "FFS claims"), num=c(sum(FFS_count), sum(HMO_count), 56763, 59361)), aes(x=source, y=num, fill=source))+ 
geom_bar(position = "dodge",stat="identity")+
   xlab("")+ylab("No. unique ids")+
 #  ylim(c(0,1))+
   theme(axis.text.x = element_text(angle =0, vjust = 0.5, size=10))+
   ggtitle("year 2019")+
   geom_text(aes(label=num), position=position_dodge(width=0.9), vjust=-0.25)+  # add numbers over bars 
  #geom_signif(annotations ="p=0.074", y_position = 0.52  ,xmin=1, xmax=2, size=1, vjust=-0.5, textsize=3.5, color="black")+ # textsize is for text over the bar; size is for width of lines 
  # geom_signif(annotations ="p<0.001", y_position = 0.59  ,xmin=1, xmax=3, size=1, vjust=-0.5, textsize=3.5, color="red")+
  #geom_signif(annotations ="p<0.001", y_position = 0.86  ,xmin=1, xmax=4, size=1, vjust=-0.5, textsize=3.5, color="red")+
   theme(plot.title = element_text(hjust = 0.5, size=15))+  #center the title
  theme(legend.title=element_blank())+
  theme(legend.position = "none")
```
* FFS claim rate: 59361/4656439=1.27%; HMO claim rate: 56763/9775082=0.58%


## 2. compare 5 categories in dental procedure HMO vs FFS 


```{r, message=F, warning=F}
data_year2019=readRDS("C:\\Shengtong\\Research\\OralPublicHealth\\Dental_Emergency_visit\\CWFluoridationAnalysis\\data_year2019.rds")
#colnames(data_year2019$hmo_claims_2019)
#unique(data_year2019$hmo_claims_2019$PROCEDURE_CODE)
```



```{r, message=F, warning=F}
hmo_code=unique(data_year2019$hmo_claims_2019$PROCEDURE_CODE)
ffs_code=unique(data_year2019$ffs_claims_2019$PROCEDURE_CODE)
hmo_ffs_code=intersect(hmo_code, ffs_code)

num_id_hmo=length(unique(data_year2019$hmo_claims_2019$RECIPIENT_ID))
num_id_ffs=length(unique(data_year2019$ffs_claims_2019$RECIPIENT_ID))

calculate_prop=function(code)
{
  hmo_prop=length(unique(data_year2019$hmo_claims_2019 %>% filter(PROCEDURE_CODE==code) %>% select(RECIPIENT_ID) %>% pull))/num_id_hmo
  ffs_prop=length(unique(data_year2019$ffs_claims_2019 %>% filter(PROCEDURE_CODE==code) %>% select(RECIPIENT_ID) %>% pull))/num_id_ffs
  return(list(hmo_prop, ffs_prop))
}

plot_figure=function(data, title, angle=0, num_size=3)
{
  
  fig=ggplot(data, aes(x=code, y=prop,fill=plan)) +geom_bar(position = "dodge", stat="identity")+
 # ylim(c(0, 0.7))+
  ylab("% unique ids")+
  xlab("")+
  theme(legend.position="bottom")+
  theme(axis.text.x = element_text(angle = angle, vjust = 0.5, size=11))+
  geom_text(aes(label=prop*100), position=position_dodge(width=0.9), vjust=-0.25, size=num_size)+  # add numbers over bars
  ggtitle(title)+
  theme(plot.title = element_text(hjust = 0.5, size=15))  #center the title
  
  return(fig)
}

```

### Diagnostic All CDT codes in the range of D0100 to D0999

```{r, message=F, warning=F}
diagnostic_code=hmo_ffs_code[substr(hmo_ffs_code, start = 1, stop = 2)=="D0"]
y=sapply(diagnostic_code,calculate_prop)


fig=plot_figure(data.frame(prop=round(unlist(y),3), code=rep(diagnostic_code, each=2), plan=rep(c("hmo", "ffs"), length(diagnostic_code))), title="Diagnostic All CDT codes", angle=30)
fig


data.frame(prop=round(unlist(y),3), code=rep(diagnostic_code, each=2), plan=rep(c("hmo", "ffs"), length(diagnostic_code)))%>%
datatable(extensions = 'Buttons',
            options = list(dom = 'Blfrtip',
                           buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
                           lengthMenu = list(c(10,25,50,-1),
                                             c(10,25,50,"All"))))
```

### Preventive procedures- D1110,  D1120, D1206, D1203, D1351, D4341 and D4342

```{r, message=F, warning=F}
preventive_code=c("D1110",  "D1120", "D1206", "D1203", "D1351", "D4341", "D4342")
y=sapply(preventive_code,calculate_prop)


fig=plot_figure(data.frame(prop=round(unlist(y),3), code=rep(preventive_code, each=2), plan=rep(c("hmo", "ffs"), length(preventive_code))), title="Preventive procedures", angle=0)
fig
data.frame(prop=round(unlist(y),3), code=rep(preventive_code, each=2), plan=rep(c("hmo", "ffs"), length(preventive_code)))%>%
datatable(extensions = 'Buttons',
            options = list(dom = 'Blfrtip',
                           buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
                           lengthMenu = list(c(10,25,50,-1),
                                             c(10,25,50,"All"))))
```

* no individual for code "D1203


### Simple restorative D2000 to D2430

```{r, message=F, warning=F}
restorative_code=sort(hmo_ffs_code[substr(hmo_ffs_code, start = 1, stop = 2)=="D2"])[1:13]
y=sapply(restorative_code,calculate_prop)


fig=plot_figure(data.frame(prop=round(unlist(y),4), code=rep(restorative_code, each=2), plan=rep(c("hmo", "ffs"), length(restorative_code))), title="Simple restorative", angle=10)
fig

data.frame(prop=round(unlist(y),4), code=rep(restorative_code, each=2), plan=rep(c("hmo", "ffs"), length(restorative_code)))%>%
datatable(extensions = 'Buttons',
            options = list(dom = 'Blfrtip',
                           buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
                           lengthMenu = list(c(10,25,50,-1),
                                             c(10,25,50,"All"))))
```


### complex restorative procedures (D2510 to D2999)

```{r, message=F, warning=F}
restorative_code=sort(hmo_ffs_code[substr(hmo_ffs_code, start = 1, stop = 2)=="D2"])[14:26]
y=sapply(restorative_code,calculate_prop)


fig=plot_figure(data.frame(prop=round(unlist(y),5), code=rep(restorative_code, each=2), plan=rep(c("hmo", "ffs"), length(restorative_code))), title="Simple restorative", angle=10, num_size=2)
fig

data.frame(prop=round(unlist(y),5), code=rep(restorative_code, each=2), plan=rep(c("hmo", "ffs"), length(restorative_code)))%>%
datatable(extensions = 'Buttons',
            options = list(dom = 'Blfrtip',
                           buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
                           lengthMenu = list(c(10,25,50,-1),
                                             c(10,25,50,"All"))))
```



### Endodontic procedures included D3000-D3999

```{r, message=F, warning=F}
Endodontic_code=sort(hmo_ffs_code[substr(hmo_ffs_code, start = 1, stop = 2)=="D3"])
y=sapply(Endodontic_code,calculate_prop)


fig=plot_figure(data.frame(prop=round(unlist(y),5), code=rep(Endodontic_code, each=2), plan=rep(c("hmo", "ffs"), length(Endodontic_code))), title="Endodontic", angle=0, num_size=3)
fig

data.frame(prop=round(unlist(y),5), code=rep(Endodontic_code, each=2), plan=rep(c("hmo", "ffs"), length(Endodontic_code)))%>%
datatable(extensions = 'Buttons',
            options = list(dom = 'Blfrtip',
                           buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
                           lengthMenu = list(c(10,25,50,-1),
                                             c(10,25,50,"All"))))
```


### extraction/surgical procedures- D7111, D7140, D7210

```{r, message=F, warning=F}
extra_surg_code=c("D7111", "D7140", "D7210")
y=sapply(extra_surg_code,calculate_prop)


fig=plot_figure(data.frame(prop=round(unlist(y),3), code=rep(extra_surg_code, each=2), plan=rep(c("hmo", "ffs"), length(extra_surg_code))), title="extraction/surgical procedures", angle=0, num_size=4)
fig

data.frame(prop=round(unlist(y),3), code=rep(extra_surg_code, each=2), plan=rep(c("hmo", "ffs"), length(extra_surg_code)))%>%
datatable(extensions = 'Buttons',
            options = list(dom = 'Blfrtip',
                           buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
                           lengthMenu = list(c(10,25,50,-1),
                                             c(10,25,50,"All"))))
```